Ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2018-01-22 manuscript/usses_water_ms.Rmd
trivia

We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA-ARS Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
The plant community is dominated by the shrub \emph{Artimesia tripartita}, \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
During the period of our experiment (2011 -- 2016), mean annual precipitation was 250 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
In 2007, we relocated 14 of the original quadrats in permenanent livestock exclosures, \new{which we used} as control plots (i.e. ambient precipitation) in the experiment described below.
\new{We used the relocated plots as our controls because collecting demographic data is time consuming and it was already being collected in these plots for other studies.}

@@

Community composition was similar among treatments.
The multidimensional space of community composition overlapped among treatments in all years and was equally dispersed in all years (Table 2; Fig. 4).
Community composition was also remarkably stable over time, with no evidence of divergence among treatments (Table 2; Fig. 4).
\new{There were also no changes in the dominant species over time in any treatment (Appendix 1).
Analyzing annual species on their own produced similar results.
There is some evidence that the annual community changed in response to treatment in two years (Fig. A1-6), but these responses appear to come from very rare species since our analysis weights common and rare species equally (Fig. A1-4).
By definition, these rare species contribute little to ANPP and thus any compositional changes have little influence on ecosystem functional response.}

@@

\begin{description}
\item [Appendix 1.] \new{Additional information on plant community composition and dynamics, Table A1-1, Table A1-2, Table A1-3, Fig. A1-1, Fig. A1-2, Fig. A1-3, Fig. A1-4, Fig. A1-5, and Fig. A1-6.}
\item [Appendix 2.] \new{Additional details on soil moisture modeling with SOILWAT and Fig. A2-1.}
\item [Appendix 3.] Additional methods and information on estimating aboveground net primary productivity, Table A3-1, and Table A3-2.
\item [Appendix 4.] Details of the hierarchical Bayesian model, Fig. A4-1, Fig. A4-2, and Fig. A4-3.
\item [Appendix 5.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A5-1.}
\item [Appendix 6.] Details on analysis of precipitation historical range of variability and Fig. A6-1.
\end{description}

>>>>

We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA-ARS Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
The plant community is dominated by the shrub \emph{Artemisia tripartita}, \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
During the period of our experiment (2011 -- 2016), mean annual precipitation was 250 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
In 2007, we relocated 14 of the original quadrats in permenanent livestock exclosures, \new{which we used} as control plots (i.e. ambient precipitation) in the experiment described below.
\new{We used the relocated plots as our controls because collecting demographic data is time consuming and it was already being collected in these plots for other studies.}

@@

Community composition was similar among treatments.
The multidimensional space of community composition overlapped among treatments in all years and was equally dispersed in all years (Table 2; Fig. 4).
Community composition was also remarkably stable over time, with no evidence of divergence among treatments (Table 2; Fig. 4).
\new{There were also no changes in the dominant species over time in any treatment (Appendix 1).
Analyzing annual species on their own produced similar results.
There is some evidence that the annual community changed in response to treatment in two years (Table A1-1; Fig. A1-5), but these responses appear to come from very rare species since our analysis weights common and rare species equally (Fig. A1-4).
By definition, these rare species contribute little to ANPP and thus any compositional changes have little influence on ecosystem functional response.}

@@

\begin{description}
\item [Appendix 1.] \new{Additional information on plant community composition and dynamics, Table A1-1, Fig. A1-1, Fig. A1-2, Fig. A1-3, Fig. A1-4, and Fig. A1-5.}
\item [Appendix 2.] \new{Additional details on soil moisture modeling with SOILWAT and Fig. A2-1.}
\item [Appendix 3.] Additional methods and information on estimating aboveground net primary productivity, Table A3-1, and Table A3-2.
\item [Appendix 4.] Details of the hierarchical Bayesian model, Fig. A4-1, Fig. A4-2, and Fig. A4-3.
\item [Appendix 5.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A5-1.}
\item [Appendix 6.] Details on analysis of precipitation historical range of variability and Fig. A6-1.
\end{description}
